---
title: "Untitled"
author: "Tegan Lohman"
date: "August 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(gridExtra)

getwd()

```



```{r}

# Load data, clean up table, check types
pisadata <- read.table('sample2.csv', header = TRUE, sep = ",")
pisadict <- read.csv('pisadict2012.csv', header = FALSE)

ncol(pisadata)
nrow(pisadict)

header <- t(pisadict)[1,]
names(pisadata) <- header

summary(pisadata)
View(pisadata[502:506])

str(pisadata)
```

```{r}

colnames(pisadata)[colnames(pisadata)=="Country code 3-character"] <- "Country"

country_groups <- group_by(pisadata, Country)

ggplot(data = country_groups, aes(factor(Country))) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Most countries have about 50 students represented in this sample, however a few are heavily over or under-represented. The United States is represented, as are several individual states. I wonder if all states have individual data in the complete data set.
```{r}
avg_scores_by_country <- aggregate(pisadata[, c(502, 542, 547)],
                                   list(pisadata$Country), mean)
names(avg_scores_by_country) <- c('country', 'math', 'reading', 'science')
ggplot(avg_scores_by_country, ) +
  geom_point(aes(country, math), color = 'blue') +
  geom_point(aes(country, reading), color = 'red') +
  geom_point(aes(country, science), color = 'green') +
  the?me(axis.text.x = element_text(angle = 90, hjust = 1))
```
The scores for math, reading, and science tend to fall within a narrow range by country. In other words, countries tend to either do well in everything, or poorly in everything.
```{r}
scores_and_ses <- read.csv('selected_columns_sample.csv', colClasses = c('factor', 'numeric', 'numeric', 'numeric', 'numeric'))

View(scores_and_ses$escs)

ggplot(scores_and_ses) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.2) +
   geom_point(aes(escs, reading), color = 'red', alpha = 0.2) +
   geom_point(aes(escs, science), color = 'green', alpha = 0.2) 

```
Internationally, students from families who are better off financially score better on the math, reading, and science assessments. Does this correlation vary by country?
```{r}
United_States <- subset(scores_and_ses, country == 'United States of America')
China <- subset(scores_and_ses, country == 'Macao-China')
Germany <- subset(scores_and_ses, country == 'Germany')
Brazil <- subset(scores_and_ses, country == 'Brazil')

p1 <- ggplot(United_States) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.2) +
  geom_point(aes(escs, reading), color = 'red', alpha = 0.2) +
  geom_point(aes(escs, science), color = 'green', alpha = 0.2) +
  ggtitle("United States")

p2 <- ggplot(Brazil) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.2) +
  geom_point(aes(escs, reading), color = 'red', alpha = 0.2) +
  geom_point(aes(escs, science), color = 'green', alpha = 0.2) +
  ggtitle("Brazil")

p3 <- ggplot(China) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.2) +
  geom_point(aes(escs, reading), color = 'red', alpha = 0.2) +
  geom_point(aes(escs, science), color = 'green', alpha = 0.2) +
  ggtitle("China")

p4 <- ggplot(Germany) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.2) +
  geom_point(aes(escs, reading), color = 'red', alpha = 0.2) +
  geom_point(aes(escs, science), color = 'green', alpha = 0.2) +
  ggtitle("Germany")

grid.arrange(p1, p2, p3, p4, ncol = 2)

```
The link between socio-economic status and test scores varies by country.

```{r}

scores_and_ses_complete <- read.csv('selected_columns_complete.csv', header = TRUE, colClasses = c('factor', 'numeric', 'numeric', 'numeric', 'numeric'))

avg_scores_by_country_complete <- aggregate(scores_and_ses_complete[, 
                                                                c(2, 3,4,5), ],
                                   list(scores_and_ses_complete$country), mean,
                                   na.rm = TRUE)

names(avg_scores_by_country_complete) <- c('country', 'escs', 'math_avg', 'reading_avg', 'science_avg')

write.table(avg_scores_by_country_complete, 
            file = "avg_scores_by_country_complete.csv",
            sep = ",")

```

```{r}


ggplot(data = avg_scores_by_country_complete, aes(factor(country), math_avg)) +
  geom_bar(stat = 'identity') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(data = avg_scores_by_country_complete, aes(escs, math_avg)) +
  geom_point()

```
```{r}
United_States_complete <- subset(scores_and_ses_complete, 
                                 country == 'United States of America')

ggplot(United_States_complete) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.1) +
  ggtitle("United States")

Canada_complete <- subset(scores_and_ses_complete, 
                                 country == 'Canada')

ggplot(Canada_complete) +
  geom_point(aes(escs, math), color = 'blue', alpha = 0.1) +
  ggtitle("Canada")

x = Canada_complete$escs
y = Canada_complete$math
cor(x,y, use = 'complete.obs')
cor(United_States_complete$escs, United_States_complete$math, use = 'complete.obs')

```

```{r}

x <- quantile(United_States_complete$escs, prob = seq(0, 1, length = 5), na.rm =TRUE)
US1 <- mean(subset(United_States_complete, United_States_complete$escs <= x[1])$math)
US2 <- mean(subset(United_States_complete, United_States_complete$escs <= x[2] &
                United_States_complete$escs > x[1])$math)
US3 <- mean(subset(United_States_complete, United_States_complete$escs <= x[3] &
                United_States_complete$escs > x[2])$math)
US4 <- mean(subset(United_States_complete, United_States_complete$escs <= x[4] &
                United_States_complete$escs > x[3])$math)
US5 <- mean(subset(United_States_complete, United_States_complete$escs > x[4])$math)
y = c(US1, US2, US3, US4, US5)
df <- data_frame(x,y)

ggplot(df, aes(x,y)) +
  geom_point() +
  xlab("ESCS by quintile") +
  ylab("average math score") +
  ggtitle("United States")

write.table(df, 
            file = "US_quantiles.csv",
            sep = ",")
```

```{r}
country_groups <- group_by(scores_and_ses_complete, country)
countries <- unique(scores_and_ses_complete$country)
```


```{r}



quintiles <- data.frame(country = character(),
                 quintile = numeric(),
                 escs = numeric(),
                 math = numeric(),
                 reading = numeric(),
                 science = numeric(),
                 stringsAsFactors = FALSE)
for (c in countries)
{
  data <- subset(scores_and_ses_complete, country == c)
  quint <-  quantile(data$escs, prob = seq(0, 1, length = 6), na.rm =TRUE)
  Q0 <- subset(data, escs <= quint[1])
  row <- data.frame(c, 0.0, quint[[1]], median(Q0$math, na.rm =TRUE), 
           median(Q0$reading, na.rm =TRUE),
           median(Q0$science, na.rm =TRUE))
  names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
  #quintiles <- rbind(quintiles, row)
  Q1 <- subset(data, data$escs <= quint[2])
    row <- data.frame(c, 0.2, quint[[2]], median(Q1$math, na.rm =TRUE), 
           median(Q1$reading, na.rm =TRUE),
           median(Q1$science, na.rm =TRUE))
  names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
  quintiles <- rbind(quintiles, row)
  Q2 <- subset(data, data$escs <= quint[3] & data$escs > quint[2])
    row <- data.frame(c, 0.4, quint[[3]], median(Q2$math, na.rm =TRUE), 
           median(Q2$reading, na.rm =TRUE),
           median(Q2$science, na.rm =TRUE))
    names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
    quintiles <- rbind(quintiles, row)
  Q3 <- subset(data, data$escs <= quint[4] & data$escs > quint[3])
    row <- data.frame(c, 0.6, quint[[4]], median(Q3$math, na.rm =TRUE), 
           median(Q3$reading, na.rm =TRUE),
           median(Q3$science, na.rm =TRUE))
    names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
    quintiles <- rbind(quintiles, row)
  Q4 <- subset(data, data$escs <= quint[5] & data$escs > quint[4])
    row <- data.frame(c, 0.8, quint[[5]], median(Q4$math, na.rm =TRUE), 
           median(Q4$reading, na.rm =TRUE),
           median(Q4$science, na.rm =TRUE))
    names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
    quintiles <- rbind(quintiles, row)
  Q5 <- subset(data, data$escs > quint[5])
    row <- data.frame(c, 1.0, quint[[6]], median(Q5$math, na.rm =TRUE), 
           median(Q5$reading, na.rm =TRUE),
           median(Q5$science, na.rm =TRUE))
    names(row) <- c('country', 'prob', 'escs', 'math', 'reading', 'science')
    quintiles <- rbind(quintiles, row)

}



```

```{r}
write.table(quintiles, 
            file = "quintiles.csv",
            sep = ",")

```
```{r}

ggplot(data = subset(quintiles, country == "United States of America"), aes(escs, math)) +
  geom_point()

unique(quintiles$country)
```

